\boldmath Flavour Symmetry Embedded - GLoBES

### 2020-04-30 doc/FAS-GLoBES.tex
update code and docs

The user can use \textbf{FASE-GLoBES} to study the constraint of model parameters. Here we take tri-direct littlest seesaw as an example to present how we can adopt \textbf{FASE-GLoBES} to constrain two model parameters -- $x$ and $\eta$. 
The $\chi^2$ function is construct based on a log -likelihood ratio,
\begin{equation}\label{eq:chi-squared}
\chi^2(\vec{\theta},\xi_s,\xi_b)=2\sum_i\left(\eta_i(\vec{\theta},\xi_s,\xi_b)-n_i+n_i\ln\frac{n_i}{\eta_i(\vec{\theta},\xi_s,\xi_b)} \right)+p(\xi_s,\sigma_s)+p(\xi_b,\sigma_b)+\chi^2_{prior},
\end{equation}
where $i$ runs over the number of bins, $\eta_i{\vec{\theta},\xi_s,\xi_b}$ is the hypothesis event rate for bin i and $E_i$ is the central bin energy. The vector $\vec{\theta}$ consist test model or oscillation parameters. Here, $\vec{\theta}=(x,\eta,r,m_a)$. The parameters $\xi_s$ and $\xi_b$ are introduced to account for the systematic uncertainty of normalisation for the signal (subscript $_s$) and background (subscript $_b$) components for the event rate, and are allowed to bary in the fit as nuisance parameters. For a given hypothesized set of parmaeters $\vec{\theta}$, the event rate for bin $i$ is calculated as\\
\begin{equation}
\eta_i(\vec{\theta},\xi_s,\xi_b)=(1+\xi_s)\times n_i+(1+\xi_b)\times b_i,
\end{equation}
where $n_i$ and $b_i$ are the expected number of signal and background events in bin $i$, respectively. The nuisance parameters are constrained by the Gaussian prior $p(\xi,\sigma)=\xi^2/\sigma^2$ with corresponding uncertainties $\sigma_s$ and $\sigma_b$ for the signal and background, respectively. Finally, $\chi^2_{prior}$ is a set of Gaussian priors for hypothesis, and is expressed as Eq.~\ref{eq:prior}. 
To get the constraint of $x$ and $\eta$, we obtain the minimum of $\chi^2$ value,\vspace{0.2cm}\\
\texttt{
    glbSetProjection(free);\\
    float res0=glbChiNP(true\_values,NULL,GLB\_ALL);\\ } \vspace{0.2cm}\\
%
Then, we set two loops to get the $\Delta \chi^2$ value for different $x$ and $\eta$\vspace{0.2cm}\\
%
\texttt{     float x,eta,r,ma,dx,deta;\\
     float lower\_x,upper\_x,lower\_eta,upper\_eta;\\
     FILE* File=fopen("data/constraint\_x\_eta.dat", "w");\\
    lower\_x=-9; upper\_x=-3; lower\_eta=0.8*M\_PI; upper\_eta=2*M\_PI;\\
    dx=(upper\_x-lower\_x)/100; deta=(upper\_eta-lower\_eta)/100;\\
        glbSetProjection(projection);\\
   for (x=lower\_x;x<=upper\_x;x=x+dx)\{ \\
    for (eta=lower\_eta;eta<=upper\_eta;eta=eta+deta)\{\\
     glbSetOscParams(test\_values,x,0); glbSetOscParams(test\_values,eta,1);\\
        float res=glbChiNP(test\_values,NULL,GLB\_ALL);\\
            fprintf(File,"\%f \%f  \%f $\backslash$n",x,eta/M\_PI,res-res0);\\
                     \} fprintf(File,"$\backslash$n");\}\\
}

@@

The example code is given below.

\texttt{ \{ ...  initalize the code ...\}\\
\{ ...  register the probability engine ...\}    }

\texttt{    
    float degree   = M\_PI/180;\\
    float x\_true,eta\_true,r\_true,ma\_true;\\
    x\_true=-3.65029; eta\_true=1.13067*M\_PI; r\_true=0.511325; ma\_true=3.71199e-3;\\
    double M\_para[4],OSC\_para[6];\\
    M\_para[0]=x\_true; M\_para[1]=eta\_true; M\_para[2]=r\_true; M\_para[3]=ma\_true;\\
    MtoS(OSC\_para, M\_para);\\
    double th12\_true=OSC\_para[0];\\
    double th13\_true=OSC\_para[1];\\
    double th23\_true=OSC\_para[2];\\
    double dCP\_true=OSC\_para[3];\\
    double DM21\_true=OSC\_para[4];\\
    double DM31\_true=OSC\_para[5];\\
    glb\_params true\_values = glbAllocParams();\\
    glb\_params test\_values = glbAllocParams();\\
    glb\_params input\_errors = glbAllocParams();\\
    glb\_params centers = glbAllocParams();\\
    \\
    glbDefineParams(test\_values,x\_true,eta\_true,r\_true,ma\_true,0,0);\\
    glbSetDensityParams(test\_values,1.0,GLB\_ALL);\\
    glbDefineParams(true\_values,th12\_true,th13\_true,th23\_true,dCP\_true,DM21\_true,DM31\_true);\\
    glbSetDensityParams(true\_values,1.0,GLB\_ALL); }

@@

>>>>

The user can use \textbf{FASE-GLoBES} to study the constraint of model parameters. Here we take tri-direct littlest seesaw as an example to present how we can adopt \textbf{FASE-GLoBES} to constrain two model parameters -- $x$ and $\eta$. 
The $\chi^2$ function is constructed based on a log-likelihood ratio,
\begin{equation}\label{eq:chi-squared}
\chi^2(\vec{\theta},\xi_s,\xi_b)=2\sum_i\left(\eta_i(\vec{\theta},\xi_s,\xi_b)-n_i+n_i\ln\frac{n_i}{\eta_i(\vec{\theta},\xi_s,\xi_b)} \right)+p(\xi_s,\sigma_s)+p(\xi_b,\sigma_b)+\chi^2_{prior},
\end{equation}
where $i$ runs over the number of bins, $\eta_i{\vec{\theta},\xi_s,\xi_b}$ is the hypothesis event rate for bin i and $E_i$ is the central bin energy. The vector $\vec{\theta}$ consist test model or oscillation parameters. Here, $\vec{\theta}=(x,\eta,r,m_a)$. The parameters $\xi_s$ and $\xi_b$ are introduced to account for the systematic uncertainty of normalisation for the signal (subscript $_s$) and background (subscript $_b$) components for the event rate, and are allowed to bary in the fit as nuisance parameters. For a given hypothesized set of parmaeters $\vec{\theta}$, the event rate for bin $i$ is calculated as\\
\begin{equation}
\eta_i(\vec{\theta},\xi_s,\xi_b)=(1+\xi_s)\times n_i+(1+\xi_b)\times b_i,
\end{equation}
where $n_i$ and $b_i$ are the expected number of signal and background events in bin $i$, respectively. The nuisance parameters are constrained by the Gaussian prior $p(\xi,\sigma)=\xi^2/\sigma^2$ with corresponding uncertainties $\sigma_s$ and $\sigma_b$ for the signal and background, respectively. Finally, $\chi^2_{prior}$ is a set of Gaussian priors for hypothesis, and is expressed as Eq.~\ref{eq:prior}. 
To get the constraint of $x$ and $\eta$, we obtain the minimum of $\chi^2$ value,\vspace{0.2cm}\\
\texttt{
    glbSetProjection(free);\\
    float res0=glbChiNP(true\_values,NULL,GLB\_ALL);\\ } \vspace{0.2cm}\\
%
Then, we set two loops to get the $\Delta \chi^2$ value for different $x$ and $\eta$\vspace{0.2cm}\\
%
\texttt{     float x,eta,r,ma,dx,deta;\\
     float lower\_x,upper\_x,lower\_eta,upper\_eta;\\
     FILE* File=fopen("data/constraint\_x\_eta.dat", "w");\\
    lower\_x=-9; upper\_x=-3; lower\_eta=0.8*M\_PI; upper\_eta=2*M\_PI;\\
    dx=(upper\_x-lower\_x)/100; deta=(upper\_eta-lower\_eta)/100;\\
        glbSetProjection(projection);\\
   for (x=lower\_x;x<=upper\_x;x=x+dx)\{ \\
    for (eta=lower\_eta;eta<=upper\_eta;eta=eta+deta)\{\\
     glbSetOscParams(test\_values,x,0); glbSetOscParams(test\_values,eta,1);\\
        float res=glbChiNP(test\_values,NULL,GLB\_ALL);\\
            fprintf(File,"\%f \%f  \%f $\backslash$n",x,eta/M\_PI,res-res0);\\
                     \} fprintf(File,"$\backslash$n");\}\\
}

@@

The example code is given below.\vspace{0.2cm}\\
\texttt{ \{ ...  initalize the code ...\}\\
\{ ...  register the probability engine ...\}    }\vspace{0.2cm}\\
\texttt{    
    float degree   = M\_PI/180;\\
    float x\_true,eta\_true,r\_true,ma\_true;\\
    x\_true=-3.65029; eta\_true=1.13067*M\_PI; r\_true=0.511325; ma\_true=3.71199e-3;\\
    double M\_para[4],OSC\_para[6];\\
    M\_para[0]=x\_true; M\_para[1]=eta\_true; M\_para[2]=r\_true; M\_para[3]=ma\_true;\\
    MtoS(OSC\_para, M\_para);\\
    double th12\_true=OSC\_para[0];\\
    double th13\_true=OSC\_para[1];\\
    double th23\_true=OSC\_para[2];\\
    double dCP\_true=OSC\_para[3];\\
    double DM21\_true=OSC\_para[4];\\
    double DM31\_true=OSC\_para[5];\\
    glb\_params true\_values = glbAllocParams();\\
    glb\_params test\_values = glbAllocParams();\\
    glb\_params input\_errors = glbAllocParams();\\
    glb\_params centers = glbAllocParams();\\
    \\
    glbDefineParams(test\_values,x\_true,eta\_true,r\_true,ma\_true,0,0);\\
    glbSetDensityParams(test\_values,1.0,GLB\_ALL);\\
    glbDefineParams(true\_values,th12\_true,th13\_true,th23\_true,dCP\_true,DM21\_true,DM31\_true);\\
    glbSetDensityParams(true\_values,1.0,GLB\_ALL); }\vspace{0.2cm}\\
\texttt{ \{ ...  set up for prior ...\}}\vspace{0.2cm}\\
\texttt{
    float th23,dCP,dth23,ddCP,lower\_th23,upper\_th23,lower\_dCP,upper\_dCP;\\
    FILE* File=fopen("data/ModelTest\_th23\_dCP\_test.dat", "w");\\
    lower\_th23=40; upper\_th23=50; lower\_dCP=125; upper\_dCP=392;\\
    dth23=(upper\_th23-lower\_th23)/100; ddCP=(upper\_dCP-lower\_dCP)/100;\\
    glbSetProjection(free);\\
    for (th23=lower\_th23;th23<=upper\_th23;th23=th23+dth23)\{\\
    for (dCP=lower\_dCP;dCP<=upper\_dCP;dCP=dCP+ddCP)\{\\
    glbSetOscParams(true\_values,th23*degree,GLB\_THETA\_23);\\
    glbSetOscParams(true\_values,dCP*degree,GLB\_DELTA\_CP);\\
    PARA=STAN; \\
    glbSetOscillationParameters(true\_values);\\
    glbSetRates();\\
    PARA=MODEL;\\ 
        float res=glbChiNP(test\_values,NULL,GLB\_ALL);\\
            fprintf(File,"\%f \%f \%f $\backslash$n",th23,dCP,gsl\_cdf\_chisq\_Qinv(gsl\_cdf\_chisq\_Q(fabs(res),dof),1));\\
         \} fprintf(File,"$\backslash$n");\\
    \}
}\vspace{0.2cm}\\
\texttt{ \{ ...  destroy the pointer ...\}}\vspace{0.2cm}\\
Finally, in the code we adopt Wilk's theorem \cite{Wilks:1938dza}. When comparing nested models, the $\Delta \chi^2$ test statistics is a random variable asymptotically distributed according to the $\chi^2$-distribution with the number of degrees of freedom, which is equal to the difference in the number of free model parameters \texttt{dof}. Here the number \texttt{dof} is $2$. And that is our output: \texttt{gsl\_cdf\_chisq\_Qinv(gsl\_cdf\_chisq\_Q(fabs(res),dof),1)}.

@@

We use the same code to analysis the exclusion ability for other oscillation-parameter combinations. Then, we have the results shown in Fig.~\ref{fig:SR_2D}.
